##Peer Assesment 1 data personal movement using activity monitoring devices

### Data

The data for this assignment can be downloaded from Github repository linked attached

###Loading and preprocessing the data

Show any code that is needed to

1. Load the data 
2. Process/transform the data (if necessary) into a format suitable for your analysis

First, a directory is created to keep the data in it, afterwards the data is downloaded and unzipped and saved in the directory. First 5 rows are shown 

```{r}
if(!file.exists("./keepdata")){dir.create("./keepdata")}
URL<-"http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(URL, destfile="./keepdata/activity.zip", "curl")
unzip("./keepdata/activity.zip", exdir = "./keepdata", overwrite = TRUE)
activity <- read.csv("./keepdata/activity.csv", sep = ",", na.strings = "NA")
head(activity)
```
###What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in the dataset.
1. Make a histogram of the total number of steps taken each day
2. Calculate and report the mean and median total number of steps taken per day

Firstly, numbers of steps are summed up by day, removing missing values afterwards a historgram is created 

```{r}
totalSteps_Day <- aggregate( steps ~ date, data = activity, FUN = 'sum',na.rm = TRUE)
hist(totalSteps_Day$steps, main = 'Total steps (Day)',xlab = 'Steps per Day', col = 'green')
```

Secondly, mean and median are reported 
```{r}
mean(totalSteps_Day$steps)
median(totalSteps_Day$steps)
```
###What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

Time series: 
```{r}
avgSteps <- aggregate( steps ~ interval, data= activity, FUN = 'mean', na.rm=TRUE)
plot(steps ~ interval, data = avgSteps, 
     type = 'l',
     main = 'Average Steps per Interval', 
     xlab = 'Interval', ylab = 'Steps',
     col = 'blue')
```
Maximum number of steps
```{r}
avgSteps[ which.max(avgSteps$steps),]$interval
```


